Examining Resource Differences by Racial Group

### 2022-09-09 analysis/analysis_main.Rmd
Add ci_df to coef_table save output

save(coef_table, file=here("analysis","output","coef_table.RData"))
```

@@

coef_table_elem <- calculate_marg_means(model4_elem)
coef_table_middle <- calculate_marg_means(model4_middle)
coef_table_hs <- calculate_marg_means(model4_hs)

coef_table <- coef_table_elem %>%
  mutate(model="elementary")
coef_table <- coef_table_middle %>%
  mutate(model="middle school") %>%
  bind_rows(coef_table)
coef_table <- coef_table_hs %>%
  mutate(model="high school") %>%
  bind_rows(coef_table) %>%
  mutate(model=factor(model, levels=c("elementary","middle school",
                                      "high school")),
         mrace=factor(mrace, levels=c("Black/White",
                                      "Black/Latino",
                                      "Black/Asian",
                                      "Black/Indigenous",
                                      "White/Indigenous",
                                      "White/Latino",
                                      "White/Asian",
                                      "Latino/Asian",
                                      "Indigenous/Latino",
                                      "Indigenous/Asian")))

ggplot(coef_table, aes(x=term, y=estimate, color=multiracial,
                       ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

@@

### Models with stricter parent linking

```{r models-parent-link, warning=FALSE, fig.height=12, fig.cap="Probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level. Panels compare results depending on whether results are restricted only to children linked to a head of household or not."}
#run separate models by grade groups
model4_best_link <- update(model4, design=subset(acs_design, momrule==11 & poprule==11))

coef_table_best_link <- calculate_marg_means(model4_best_link)

coef_table <- coef_table_best_link %>%
  mutate(model="only direct links")
coef_table <- coef_table_mod4 %>%
  mutate(model="all links") %>%
  bind_rows(coef_table) %>%
  mutate(model=factor(model, levels=c("all links","only direct links")),
         mrace=factor(mrace, levels=c("Black/White",
                                      "Black/Latino",
                                      "Black/Asian",
                                      "Black/Indigenous",
                                      "White/Indigenous",
                                      "White/Latino",
                                      "White/Asian",
                                      "Latino/Asian",
                                      "Indigenous/Latino",
                                      "Indigenous/Asian")))

ggplot(coef_table, aes(x=term, y=estimate, color=multiracial,
                       ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

The results here are extraordinarily stable across both specifications.

>>>>

save(coef_table, ci_df, file=here("analysis","output","coef_table.RData"))
```

@@

results_elem <- calculate_marg_means(model4_elem)
results_middle <- calculate_marg_means(model4_middle)
results_hs <- calculate_marg_means(model4_hs)

coef_table <- results_elem$coef_table %>%
  mutate(model="elementary")
coef_table <- results_middle$coef_table %>%
  mutate(model="middle school") %>%
  bind_rows(coef_table)
coef_table <- results_hs$coef_table %>%
  mutate(model="high school") %>%
  bind_rows(coef_table) %>%
  mutate(model=factor(model, levels=c("elementary","middle school",
                                      "high school")),
         mrace=factor(mrace, levels=c("Black/White",
                                      "Black/Latino",
                                      "Black/Asian",
                                      "Black/Indigenous",
                                      "White/Indigenous",
                                      "White/Latino",
                                      "White/Asian",
                                      "Latino/Asian",
                                      "Indigenous/Latino",
                                      "Indigenous/Asian")))

ci_df <- results_elem$ci_df %>%
  mutate(model="elementary")
ci_df <- results_middle$ci_df %>%
  mutate(model="middle school") %>%
  bind_rows(ci_df)
ci_df <- results_hs$ci_df %>%
  mutate(model="high school") %>%
  bind_rows(ci_df) %>%
  mutate(model=factor(model, levels=c("elementary","middle school",
                                      "high school")),
         mrace=factor(mrace, levels=c("Black/White",
                                      "Black/Latino",
                                      "Black/Asian",
                                      "Black/Indigenous",
                                      "White/Indigenous",
                                      "White/Latino",
                                      "White/Asian",
                                      "Latino/Asian",
                                      "Indigenous/Latino",
                                      "Indigenous/Asian")))

ggplot(coef_table, aes(x=term, y=estimate, color=colors))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=lower, ymax=upper, fill=colors), 
            color=NA, alpha=0.5)+
  geom_linerange(data=ci_df, 
                 position=position_dodge2(width = 0.1), size=1, alpha=0.5,
                 aes(ymin=lower, ymax=upper, color=colors))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=color_choices)+
  scale_fill_manual(values=color_choices)+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

@@

### 2022-09-09 analysis/analysis_resources.Rmd

```{r create-cf-marg}
coefs_cf <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

coef_table_cf <- create_coef_table(coefs_cf)
```

@@

```{r counterfactual-ses, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were resources (household income, parental education, home ownership, parent and child foreign-born status, parent and child English proficiency, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
ggplot(coef_table_cf, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")

@@

```{r cf-ses-isolate, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were SES resources (household income, parental education, home ownership, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
cf_df_ses <- cf_df
cf_df_ses$foreign_born <- FALSE
cf_df_ses$foreign_born_father <- FALSE
cf_df_ses$foreign_born_mother <- FALSE
cf_df_ses$spk_eng <- "Yes"
cf_df_ses$spk_eng_father <- "Yes"
cf_df_ses$spk_eng_mother <- "Yes"

coefs_cf_ses <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df_ses, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

coef_table_cf_ses <- create_coef_table(coefs_cf_ses)

ggplot(coef_table_cf_ses, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf_ses, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

@@

coefs_cf_culture <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df_culture, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

coef_table_cf_culture <- create_coef_table(coefs_cf_culture)

ggplot(coef_table_cf_culture, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf_culture, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

>>>>

```{r create-cf-marg}
coefs_cf <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows() %>%
    mutate(multiracial=str_detect(term, "/"))

results_cf <- create_coef_table(coefs_cf)
```

@@

```{r counterfactual-ses, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were resources (household income, parental education, home ownership, parent and child foreign-born status, parent and child English proficiency, and whether parents are married)"}
ggplot(results_cf$coef_table, aes(x=term, y=estimate, color=multiracial))+
  geom_hline(aes(yintercept=estimate), color="grey")+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")

@@

```{r cf-ses-isolate, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were SES resources (household income, parental education, home ownership, and whether parents are married)."}
cf_df_ses <- cf_df
cf_df_ses$foreign_born <- FALSE
cf_df_ses$foreign_born_father <- FALSE
cf_df_ses$foreign_born_mother <- FALSE
cf_df_ses$spk_eng <- "Yes"
cf_df_ses$spk_eng_father <- "Yes"
cf_df_ses$spk_eng_mother <- "Yes"

coefs_cf_ses <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df_ses, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows() %>%
    mutate(multiracial=str_detect(term, "/"))

results_ses <- create_coef_table(coefs_cf_ses)

ggplot(results_ses$coef_table, aes(x=term, y=estimate, color=multiracial))+
  geom_hline(aes(yintercept=estimate), color="grey")+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```

@@

coefs_cf_culture <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df_culture, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows() %>%
    mutate(multiracial=str_detect(term, "/"))

results_culture <- create_coef_table(coefs_cf_culture)

ggplot(results_culture$coef_table, aes(x=term, y=estimate, color=multiracial))+
  geom_hline(aes(yintercept=estimate), color="grey")+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")
```
